Artificial ecosystem selection to deduce the mapping between microbial community structure and ecosystem function

### 2020-07-31 Manuscript/manuscript.Rmd
create new process script and add G6 data and write new Intro.

@@

## Methane flux measurements
Methane oxidation rates were determined after flushing and spiking jars to approximately 1000 ppm CH~4~. Head-space samples of 1 mL were collected from each jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for a 5-point curve. Samples were immediately injected into a <mark>???</mark> gas chromatograph fitted with an electron capture detector (company, city, state, USA) to determine the headspace CH~4~ concentration. Fluxes were calculated from a first-order exponential decay function as decay constant `k` with units<mark>???</mark>. Oxidation rates are presented as the additive inverse of `k` (i.e., `-k`) so that a more positive value represents a faster oxidation rate.

@@

To test whether there was a significant change in CH~4~ oxidation rate (k) as a response to selection, we tested the deviation between the positive and neutral selection lines over the five passages. We calculated deviation as the difference in the mean oxidation rate between the positive and neutral line within each passage with standard errors equal to the sum of the errors for each of the positive and neutral lines. We then fit an ordinary least squares regression model taking into account the standard errors of the deviations. The model was fit using the `rma` function from the `metafor` package in `r` (Viechtbauer 2010).

@@

@@

>>>>

Ecosystem functions that humans depend on are largely controlled by biodiversity in terms of the diversity and function of organisms in an ecosystem. Therefore, to understand the impact of human alterations of the environment and to manage environments for future generations requires that we understand the effects of species on ecosystem processes (Hooper et al. 2005). The vast majority of research on the connections between biodiversity and ecosystem function have been conducted in temperate grassland ecosystems evaluating the connection between manipulated plant richness and productivity. This research has demonstrated a positive, asymptotic relationship between species richness and productivity. However, it remains unclear whether this mapping is generalizable to other functions or to other taxonomic groups.

In evolutionary biology, genotype-phenotype mapping provides a way to understand the fitness effects of a genotype. While selection typically acts on phenotype, the genotypic changes are the fundamental driver of phenotypic differences and thus fitness differences (Stadler and Stephes 2005). A similar analogy can be drawn for biodiversity-ecosystem function relationships. Traits of a community are what drive community assembly, not the taxa or genotypes themselves (Inkpen et al. 2017). However, as ecologists, we wish to understand the effects of species on ecosystem function. In a similar way that evolutionary biologists want to deduce the genotype-phenoype map to understand the genetic basis of fitness, community ecologists must describe the biodiversity-ecoystem function map to understand what drives communty assembly, and ultimately, ecosystem function.

The relationship that ecologists have described between plant diversity and productivity represents a relatively simple map between taxonomy and function. The number of unique members of a functional group (i.e., plant species) is correlated with the rate of the function performed by that group. However, for microbial communities that mapping is not quite as clear. For many microbial traits, there is not a clear coupling between taxonomy and function, except for in the case of highly conserved traits involving many genes (Martiny et al. 2013). In addition, most microbial taxa have not been sampled and so we don't have a full picture of which taxa are members of a particular functional group (citation). It's much harder to identify methanotrophic prokaryotes in soil than it is to identify photosynthetic eukaryotes in a prairie. Finally, even for highly conserved traits with a putative functional marker, for example methane oxidation and the gene *pmoA*, functional group abundance is not correlated with the rate of the function performed by that group in most ecosystems (Rocca et al. 2016). Because of these differences between macroorganisms and microorganisms, it may be more useful to approach microbial biodiversity-ecosystem function in a way analogus to genotype-phenotype mapping in organisms (Morris et al. 2020).

One approach to deducing the genotype-phenotype map in organisms is through artificial selection (Fuller et al. 2005). In this approach, selection for a trait is imposed in a contrived laboratory environment and compared to laboratory environment without selection on that trait. By comparing the genetic respones between these two treatments, you can remove the environment-fitness connection that would be present in a comparative study and identify genes that are evolving in response to selection only on the trait of interest. It has been shown that whole communities can respond to artifical selection on traits of the ecosystem (Goodnight et al. 1997, Williams and Lenton 2007). And it has been proposed that this could be a suitable approach for identify communities of organisms that are capable of performing a particular function at a high rate (Swenson et al. 2000). Therefore, artificial ecosystem selection is a potentially fruitful approach to identifying the map between biodiversity and ecosystem function in microbial communities.

To identify the map between microbial community composition and ecosystem function, we imposed artificial ecosystem selection on methane oxidation rate of whole soil ecosystems. We prepared two lines of replicate ecosystems: one in which we imposed positive selection on methane oxidation by only using the top performing ecosystems to generate the next set of ecosystems and one in which we imposed neutral selection where a random set of ecosystems were chosen to generate the next set of ecosystems. We repated this process over multiple passages and then compared the resultant metacommunities in terms of their rate of ecosystem function and the composition of their communities.

With this experiment, we asked several questions. First, is there a response to selection on methane oxidation by soil ecosystems? And how strong is the response to selection? (i.e., how "heritable" is methane oxidation as a community trait in terms of the phenotypic resemblence of ecosystems to the ecosystems that generated them?) If methane oxidation is heritable, then what is the nature of the map between microbial community composition and methane oxidation? Is it the diversity of methane oxidizers, as it is for plants and productivity? Or is there a more complex map involving upstream metabolic regulation?

Things this Intro doesn't yet discuss:

1. What a more complex map might look like. For example, maybe pmoA and methanotrophy aren't correlated, but maybe understand methanogenesis or nitrogen cycling is helpful.

2. The idea of "heritability" for communities/ecosystems and why that relates to environmental vs. genetic control over ecosystem function.

@@

## Methane flux measurements
Methane oxidation rates were determined after flushing and spiking jars to approximately 1000 ppm CH~4~. Head-space samples of 1 mL were collected from each jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for a 5-point curve. Samples were immediately injected into a <mark>???</mark> gas chromatograph fitted with an electron capture detector (company, city, state, USA) to determine the headspace CH~4~ concentration. Fluxes were calculated from a first-order exponential decay function as decay constant `k`. Oxidation rates are presented as the additive inverse of `k` (i.e., `-k`) so that a more positive value represents a faster oxidation rate.

@@

To test whether there was a significant change in CH~4~ oxidation rate (k) as a response to selection, we tested the deviation between the positive and neutral selection lines over multiple passages. We calculated deviation as the difference in the mean oxidation rate between the positive and neutral treatment within each passage. The standard errors of the deviations were calculated as the sum of the standard errors of the mean for each of the positive and neutral lines. We then fit an ordinary least squares regression model taking into account the standard errors of the deviations. The model was fit using the `rma` function from the `metafor` package in `r` (Viechtbauer 2010).

@@

ggsave('new_select_fig.pdf', width = 4, height = 3)

@@

ggsave('new_herit_fig.pdf', width = 5, height = 4)
